---
title: "Homework 1"
subtitle: "DATA 202 - Alexander - Spring 2024"
format: 
  html: 
    self-contained: true
    self-contained-math: true
editor: visual
---

Please submit **Homework 1** responses as a `.pdf` file on Canvas [here](https://canvas.howard.edu/courses/46792).

### Exercise 1.1

The [USPHS Syphilis Study at Tuskegee](https://www.cdc.gov/tuskegee/timeline.htm) was one of many historical cases of scientific racism and unethical practices enacted by the U.S. government. Was the USPHS Syphilis Study at Tuskegee an *experimental* or *observational* study? Explain your reasoning.

### Exercise 1.2

Conduct a web search and literature review to identify two (2) other cases where data and/or information was collected using unethical practices. Provide a brief explanation of each case and its significance to understanding ethics. Include a full citation of all sources.

### Exercise 1.3

There are many definitions of statistics. What is the definition of statistics that has been used in our course lectures? Conduct a web search and find two (2) alternative definitions of statistics. Include a full citation of all sources.

### Exercise 1.4

Based on your current understanding of statistics, what should it mean to be *critical* in the context of statistics? Explain your thinking. Include a full citation of all sources.

### Exercise 1.5

You have been tasked with identifying data for a new study on [homelessness](https://www.ohchr.org/en/special-procedures/sr-housing/homelessness-and-human-rights) and [housing insecurity](https://www.urban.org/sites/default/files/publication/101608/improving_measures_of_housing_insecurity.pdf) in your local area. Identify and describe two (2) of each variable type that could be collected for this study: `nominal`, `ordinal`, `discrete`, `continuous`. That is, identify and fully describe two nominal, two ordinal, two discrete, and two continuous variables that could be used to gather insights about homelessness and housing insecurity. Your descriptions should be detailed.

### Exercise 1.6

Your local city council plans to conduct a 2024 census of the local homeless population. You have been hired as a data analyst to identify priorities for the project. This census is part of a broader effort to understand the issues experienced during homelessness and to find ways to mitigate issues given the recent increase in the homeless population. What are some potential *ethical issues* that could arise with the councils' plans to conduct a local homeless census? In the city's plan to collect data on those experiencing homelessness, should *informed consent* be obtained per the IRB? If so, why and how?

### Exercise 1.7

Describe the contents of the data set below and what the values most likely represent.

```{r}
#| output: true
#| echo: false
#| warning: false
# library(tidyverse)
library(critstats)
head(world_pop)
tail(world_pop)
```

### Exercise 1.8

```{r}
#| output: false
sum(1:51)
```

-   What is the meaning of the code chunk `sum(1:51)`?
-   What is the numerical output?

### Exercise 1.9

The population of five countries is listed in a data set using computational scientific notation.

Numerically expand the population for each country.

-   Country 1: 2.06139E8
-   Country 2: 8.9561E7
-   Country 3: 2.77E7
-   Country 4: 2.72815E5
-   Country 5: 6.077E3

### Exercise 1.10

Describe the error in the following attempt to construct a data frame.

```{r}
#| output: false
#| echo: true
#| warning: false
vec1 <- c(1, 2, 3, 4)
vec2 <- c("a","b","c","d")
vec3 <- data.frame(T, F, F, T)
df <- data.frame(vec1, vec2, vec3)
```
