---
title: "Homework 3"
subtitle: "DATA 202 - Alexander - Spring 2024"
format: 
  html: 
    self-contained: true
    self-contained-math: true
editor: visual
---

Please submit **Homework 3** responses as a `.pdf` file on Canvas [here](https://canvas.howard.edu/courses/46792).

### Exercise 1.1

```{r}
#| echo: false
#| output: false
#| warning: false
## Sample outline of a bivariate regression analysis

# Generate sample data
set.seed(203)
x <- rnorm(75)
y <- 9 - 6*x + rnorm(75)
```

Is the relationship between the $x$ and $y$ variables in the below model significant?

If so, explain. If not, explain why.

```{r}
model <- lm(y ~ x)
summary(model)
```

### Exercise 1.2

Examine the plot below. Estimate the correlation coefficient for the plot.

```{r}
plot(x, y)
```

### Exercise 1.3

```{r}
#| echo: false
#| output: false
#| warning: false
# Generate sample data
set.seed(11)
age <- runif(n = 100, min = 18, max = 65)
income <- runif(n = 100, min = 20000, max = 152500)
```

Examine the plot below. Estimate the correlation coefficient for the plot.

Based on your estimate, should we move forward with our analysis? If so, why? If no, why not?

```{r}
plot(age, income)
```

### Exercise 1.4

```{r}
#| echo: false
#| output: false
#| warning: false
id <- c(1:100)
capacity = rpois(n = 100, lambda = 7500)
funding = rnorm(n = 100, mean = 87000, sd = 15000)
```

In a few sentences, summarize the relationship between the variables based on the output.

Is there a significant relationship?

```{r}
model2 <- lm(funding ~ capacity)
summary(model2)
plot(capacity, funding)
abline(model2, col="blue")
```

### Exercise 1.5

Using the model outlined above and the plot shown below, explain the function of a residual plot.

Does the residual plot represent a "healthy" or "problematic" pattern?

```{r}
resids2 <- residuals(model2)
plot(funding, resids2)
```
